---
title: "Assignment 7: Web Technologies"
author: "Desiree Thomas, Denise Atherley, Kiera Griffiths"
format: html
editor: visual
---

## Approach: HTML and JSON

For this assignment we have chosen three philosophy books. Since not all group members are familiar with JSON and HTML, we will manually create these structures. HTML will use the table structure and JSON will focus mostly on the data types. The JSON will handle the nesting since there are multiple authors listed for the same book; this will be held within an array. The files will be saved with .html and .json. To load, we will load the rvest package and use read_html() to read the file and in the next step we will extract the table components. We will use the jsonlite library for the JSON file and use fromJSON to read the file.

Anticipated challenges may include structuring challenges and gaining an understanding of HTML and JSON structure since they are being manually written. It is important that we follow and maintain the correct syntax to avoid issues and bugs later on. The output when viewing the JSON and HTML will also look different between the two. We will host our files on GitHub and use the RAW version for reproducibility.

#Load the HTML data into R dataframe

```{r}
library(tidyverse)
library(rvest)
library(jsonlite)
library(knitr)
```

#Load the HTML data frame and displays the table

```{r}
raw_html <- read_html("https://raw.githubusercontent.com/desithomas/607-7/refs/heads/data-loading/books.html")

#Extract the table and place it into a dataframe
html_df <- raw_html %>%
  
#Looks for the table element; it must be in a string
html_element("table") %>%

html_table()

#displayed contents in Knitr table to check if identical
kable(html_df, caption = "Book data from HTML file")
```

#Load the JSON data frame and displays the table

```{r}
json_df <-fromJSON("https://raw.githubusercontent.com/desithomas/607-7/refs/heads/data-loading/books.json")

#displayed table contents using Knitr to check if identical
kable(json_df, caption = "Book data from the JSON file")
```

## Analysis: Are the two tables identical?

We use the identical function to check if the objects are 100% clones. If they are not identical it will render as False.

```{r}
# Checking if tables are identical with identical function below

identical(html_df, json_df)
```
In order to know what specific differences exist for these tables, we will use the all.equal function. It can identify if the data is the same while ignoring minor attribute differences like table indexing.

```{r}
# Using this function to identify the specific differences between the tables

all.equal(html_df, json_df)
```

```{r}
# 1. Standardize JSON: Flatten the author list and ensure year is character
json_df_final <- json_df %>%
  mutate(
    # Flattens the list of authors into a single comma-separated string [cite: 4, 18]
    Authors = map_chr(Authors, ~ paste(.x, collapse = ", ")), 
    # Ensures the year column is a character type 
    year = as.character(year)
  )

# 2. Standardize HTML: Ensure year is character to match JSON 
html_df_final <- html_df %>%
  mutate(
    year = as.character(year)
  )

# 3. Remove any metadata/attributes that differ (like row names or class types) 
# Converting to tibble ensures both have the same base class
json_df_final <- as_tibble(json_df_final)
html_df_final <- as_tibble(html_df_final)

# 4. Final verification [cite: 14]
is_identical <- identical(html_df_final, json_df_final)

# Output results
print(paste("Are the tables identical now?", is_identical))

# Displaying both to confirm visual parity [cite: 13]
kable(html_df_final, caption = "Cleaned HTML Table")
kable(json_df_final, caption = "Cleaned JSON Table")
```


